Analysis of Global Crises and Their Impact on Interational Well-Being

### 2024-11-15 Draft.Rmd
removed extra graphs added insight 2 summary stats started intro and abstract started data change desc for insight 2

# Instructions

You are encouraged to remove this instruction section prior to submission.

It is recommended that you follow the structure of this template. The text is all placeholder - you are free to change any/all wording as you please, but it is very helpful for the grading process if you keep the same structure. Anything in \<<double angle brackets>\> definitely needs to be changed, but you are free to change any/all sentences!

Note that all of the code is *hidden* by default. This file will be graded based on the insights, not the code.

You will only submit the PDF version of this document. To knit to PDF, you'll need to run `install.packages("tinytex")` in the console, followed by `tinytex::install_tinytex()` (DO NOT PUT THESE COMMANDS IN AN RMD FILE!!!). If you encounter errors in "Knit to PDF", you can "knit to html" and then print the html file to PDF using your operating system's PDF view (e.g. Adobe Acrobat). Only standalone PDF files will be accepted by MyLS.

# Purpose
This study aims to

@@

@@

@@

## \<\<Cyclones Data\>\>
```{r load_data1}
cyclones_data <- read_parquet("cyclones_data.parquet")
world <- ne_countries(scale = "medium", returnclass = "sf")

@@

@@

A report conducted by the [National Librarbty of Medicine](https://pmc.ncbi.nlm.nih.gov/articles/PMC8079387/) highlights a strong correlation between COVID-19 cases and proximity to the equator, suggesting that certain geographic and environmental factors may influence virus spread and outbreak. This observation signifies a trend where regions near the equator are subject to unique vulnerabilities, not only experiencing increased COVID-19 cases but also facing intensified cyclone activity (as shown in the figure above), volcanic activity, elevated ocean currents, seasonal flooding, and the presence of natural carbon sinks.

@@

```{r}
#Social support & life ladder

@@

```{r insight3, fig.height=4, fig.width=6, fig.cap="This is a figure caption that you will need to change in order to get good marks in the visualization rubric items."}

covid_happiness |> filter(!is.na(total_cases), !is.na(perceptions_of_corruption)) |>
  ggplot(mapping = aes(x = perceptions_of_corruption, y = total_cases, color = continent)) + geom_boxplot() + facet_wrap(~continent) + labs(title = "Perception of Corruption and COVID cases by continent", subtitle = "Weight of COVID cases repective to perception of corruption", x = "Perception of Corruption (0 - 1)", y = "Total Cases", caption = "Source: covid.ourworldindata.org\nworldhappiness.report") + theme_minimal() + theme(legend.position = "None")

```{r insight3, fig.height=4, fig.width=6, fig.cap="This is a figure caption that you will need to change in order to get good marks in the visualization rubric items."}

>>>>

# Purpose
This study aims to investigate the relationship between freedom of citizens (i.e. freedom of choices, a just government) and quality of life for citizens. The more freedom for citizens allows for better quality of life.

@@

This paper will aim to show as freedom of citizens increase and citizen's quality of life improves. The relationship we look for is a positive correlation. First, we explore the relationship of social support and life expectancy. We observe more freedom allows for better opportunities for social support.

@@

Quality of life determines happiness which is impacted by many factors.

@@

## \<\<Cyclones Data\>\>

```{r cyclones_data}
cyclones_data <- read_parquet("cyclones_data.parquet")
world <- ne_countries(scale = "medium", returnclass = "sf")

@@

For our second insight we needed to join the COVID_2020 data along with the happiness data, to accomplish this we applied the "inner_join()" function with the argument "by = continent'". This allo

@@

A report conducted by the [National Library of Medicine](https://pmc.ncbi.nlm.nih.gov/articles/PMC8079387/) highlights a strong correlation between COVID-19 cases and proximity to the equator, suggesting that certain geographic and environmental factors may influence virus spread and outbreak. This observation signifies a trend where regions near the equator are subject to unique vulnerabilities, not only experiencing increased COVID-19 cases but also facing intensified cyclone activity (as shown in the figure above), volcanic activity, elevated ocean currents, seasonal flooding, and the presence of natural carbon sinks.

@@

```{r insight2, fig.height=4, fig.width=6, fig.cap="Boxplot of perception of corruption vs total COVID cases by continent"}

covid_happiness |> filter(!is.na(total_cases), !is.na(perceptions_of_corruption), total_cases <= 20000000) |>
  ggplot(mapping = aes(x = perceptions_of_corruption, y = total_cases, color = continent)) + geom_point() + geom_smooth(method = "lm", se = FALSE) + facet_wrap(~continent) + labs(title = "Perception of Corruption and COVID cases by continent", subtitle = "Weight of COVID cases respective to perception of corruption", x = "Perception of Corruption (0 - 1)", y = "Total Cases", caption = "Source: covid.ourworldindata.org\nworldhappiness.report") + theme_minimal() + theme(legend.position = "None")

```{r summary_stats}
library(gtsummary)
library(gapminder)

insight_2_sum_stats <- covid_happiness |> group_by(continent) |> select(continent, perceptions_of_corruption, total_cases) |> 
  tbl_summary(by = continent)
insight_2_sum_stats
```

```{r insight3, fig.height=4, fig.width=6, fig.cap="This is a figure caption that you will need to change in order to get good marks in the visualization rubric items."}
happiness <- happiness|>
  left_join(
    rnaturalearth::ne_countries(scale = "medium", returnclass = "sf") |>
      transmute(
        country = name_long,
        distance_to_equator_km = geosphere::distGeo(
          cbind(st_coordinates(st_centroid(geometry))[,1], st_coordinates(st_centroid(geometry))[,2]),
          cbind(st_coordinates(st_centroid(geometry))[,1], 0)
        ) / 1000  # Convert to km
      ),
    by = "country"
  )

```{r insight2, fig.cap=""}
#Social support & life ladder

@@
